BIP-Predictor

### 2025-08-16 Archive IWH Simon paper/16.08.2025/paper.tex
up

Let $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ denote the M-SSA predictor of HP-GDP at time $t$ for forecast horizon $h$. A GDP forecast can be obtained from the following regression specification\footnote{By default, the M-SSA criterion \eqref{eq:gen_stat_x} generates a standardized predictor and the regression coefficients are calibrated to match the level and scale of GDP.}
\begin{equation}
\textrm{GDP}_{t+shift}=\beta_0(shift,h)+\beta_1(shift,h) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h) + \epsilon_t^{shift,h}.
\label{eq:GDP_M-SSA}
\end{equation}
%The purposes of the regression are twofold: first, the parameters enable affine transformations of the left-hand side variable, so that the standardized target $\textrm{GDP}_{t+shift}$ can be replaced by the original log-returns of GDP; second, the scale adjustment by $\beta_1(shift,h)$ compensates for the deflation (zero-shrinkage) of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, reflecting greater uncertainty as the forecast horizon $h$ extends. This simple transformation allows $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ to serve as a predictor for $\textrm{GDP}_{t+shift}$ even when $h\neq shift$. This flexibility enables leveraging the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, illustrated in Figure \eqref{multivar_vs_univar}, when $shift<h$.\\
The corresponding out-of-sample predictor $\widehat{\textrm{GDP}}_{t+shift}(h)$ is given by: 
\begin{equation}\label{eq:oos_pred}
\widehat{\textrm{GDP}}_{t+shift}(h):=\hat{\beta}_0(shift,h,t)+\hat{\beta}_1(shift,h,t) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h),
\end{equation}
where the regression parameters $\hat{\beta}_0(shift,h,t),\hat{\beta}_1(shift,h,t)$ are estimated recursively using data up to time point $t-1$ (expanding window). Since the M-SSA predictor is based on data up to Q4-2007, $\widehat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all $shift\geq 0$ and all $t$ larger than Q4-2007.\footnote{Equations \eqref{eq:GDP_M-SSA} and \eqref{eq:oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.} 
The associated out-of-sample forecast error is defined as
\begin{equation}\label{eq:oosfe}
\hat{\epsilon}_t^{shift,h}:={\textrm{GDP}}_{t+shift}-\widehat{\textrm{GDP}}_{t+shift}(h).
\end{equation}
%We now assume that the regression parameters $\hat{\beta}_0(shift,h,t)$ and$,\hat{\beta}_1(shift,h,t)$ in Equation \eqref{oos_pred} are up-dated quarterly. Note that 
In the empirical implementation, the M-SSA forecasts are held fixed, while the regression parameters $\hat{\beta}_0(\cdot)$ and $\hat{\beta}_1(\cdot)$ are updated each quarter using an expanding window beginning in 2008Q1. This procedure yields a sequence of optimally weighted M-SSA forecasts that adapt over time to evolving GDP dynamics.

@@

% Obviously, predicting HP-GDP (see Table \eqref{p_val_hpbip_wc}) is less challenging than predicting GDP (see Table \eqref{p_val_wc}), due to smoothness (autocorrelation) of the target when compared to GDP. 
Our results can be summarized as follows. When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead (Table \ref{tab:rRMSE_mSSA_direct_mean_without_covid8} in the appendix). %\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.}.
In contrast, the GDP predictors $\widehat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5 quarters ahead, see Table \ref{tab:rRMSE_mSSA_comp_mean_wc}. Additionally, $\widehat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$ (Table \ref{tab:rRMSE_mSSA_comp_direct_without_covid7}). Statistical significance of the GDP predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead. Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors, see Figures \ref{fig:multivar_vs_univar} and \ref{fig:mssa_hpc_financial_crisis}, in relation to forecast performance.

%???check whether rMSE is decreasing towards sample end (because initial estimate towards 2007 is unreliable)???

\begin{table}[ht]

\caption{Relative RMSE of M-SSA GDP Predictors vs. Mean Benchmark.\\
Out-of-sample rRMSEs for various shifts and horizons, excluding the pandemic.\\
Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$; $h$ refers to the forecast horizon for which $\textrm{M-SSA}_t^{\textrm{GDP}}(h)$ is optimized. 
Values $< 1$ indicate superior predictions from Eq.~ \eqref{eq:oos_pred}. Significance refers to the hypothesis $\alpha_1=0$ in Eq.~\eqref{eq:eq_gdp_pred}.
} 
\label{tab:rRMSE_mSSA_comp_mean_wc}
\centering
\begin{tabular}{llllllll}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.975 & $0.941^{*}$ & $0.886^{*}$ &$ 0.839^{**} $& $0.843^{**}$ &$ 0.897^{**} $& 0.965 \\ 
  Shift=1 & 1.085 & 1.068 & 1.011 & $0.927^{*} $& $0.871^{**}$ & $0.875^{**}$ & $0.919^{**}$ \\ 
  Shift=2 & 1.076 & 1.089 & 1.074 & 1.011 & $0.938^{*}$ &$ 0.902^{**} $& $0.909^{**} $\\ 
  Shift=3 & 1.037 & 1.069 & 1.090 & 1.059 & 0.992 & $0.943^* $&$ 0.930^*$ \\ 
  Shift=4 & 1.013 & 1.058 & 1.101 & 1.097 & 1.037 & 0.967 &$ 0.925^*$ \\ 
  Shift=5 & 1.016 & 1.062 & 1.106 & 1.115 & 1.078 & 1.015 & $0.955^*$ \\  
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Relative RMSEs of M-SSA GDP Predictors vs. Direct Forecasts.\\
Out-of-sample evaluation starting in Q1-2008, without pandemic. Values below 1 indicate improvement over direct forecasting.
\label{tab:rRMSE_mSSA_comp_direct_without_covid7}}
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.131 & 1.092 & 1.028 & 0.973 & 0.977 & 1.040 & 1.119 \\ 
  Shift=1 & 1.295 & 1.274 & 1.206 & 1.106 & 1.039 & 1.044 & 1.097 \\ 
  Shift=2 & 1.180 & 1.194 & 1.178 & 1.109 & 1.029 & 0.990 & 0.997 \\ 
  Shift=3 & 1.029 & 1.061 & 1.082 & 1.051 & 0.984 & 0.936 & 0.923 \\ 
  Shift=4 & 1.012 & 1.057 & 1.100 & 1.096 & 1.036 & 0.967 & 0.924 \\ 
  Shift=5 & 0.988 & 1.033 & 1.076 & 1.084 & 1.048 & 0.987 & 0.929 \\  
   \hline
\end{tabular}
\end{table}

@@

\begin{table}[ht]
\caption{AUCs of predictors against GDP at various forecast horizons.} 
\label{tab:auc}
\centering
\begin{tabular}{rrrrr}
  \hline
 & Direct forecast & Direct HP forecast & M-MSE & M-SSA \\ 
  \hline
Forecast horizon 1 & 0.752 & 0.715 & 0.750 & 0.756 \\ 
  Forecast horizon 2 & 0.697 & 0.663 & 0.714 & 0.744 \\ 
  Forecast horizon 3 & 0.554 & 0.622 & 0.666 & 0.716 \\ 
  Forecast horizon 4 & 0.486 & 0.624 & 0.707 & 0.732 \\ 
  Forecast horizon 5 & 0.593 & 0.636 & 0.660 & 0.713 \\ 
   \hline
\end{tabular}

@@

For illustration, we consider a simple approach based on regressing the M-SSA predictor (of HP-GDP) on future GDP. This predictor is intuitively appealing because the future HP-GDP, i.e., the target of M-SSA, is the low-frequency component of the future GDP target. Consequently, a strong link between M-SSA and future HP-GDP also indicates a connection with GDP, even if the statistical significance is obscured by the noise inherent in GDP. Out-of-sample performance suggests that this predictor is statistically significant at horizons exceeding one year. Moreover, designs optimized for larger forecast horizons outperform the mean benchmark at all considered forecast horizons, and the predictor exceeds the performance of direct forecasts---based on unfiltered or HP filtered indicators---at horizons longer than two quarters. An analysis of hit and false alarm rates within the ROC curve confirms that filtering improves predictive performance for forecast horizons exceeding two quarters. Multivariate approaches can further capitalize on leading indicators within the cross-sectional data. Additionally, the AUC is optimized through more precise control of the sign change rate within the M-SSA framework. Finally, an examination of revision errors associated with quarterly updates of the proposed M-SSA-based GDP predictor indicates that the regression parameters, VAR model, and resulting M-SSA filters stabilize after an initial burn-in period. This stabilization enhances the model’s explainability and interpretability.

@@

\begin{table}[ht]
\caption{Significance of M-SSA GDP Component Predictor (Including Pandemic).\\
p-values for $H_0: {\alpha_1} = 0$, HAC-adjusted Wald test, Equation \eqref{eq:eq_gdp_pred}.\\ including the pandemic.} 
\label{tab:p_val1}
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.080 & 0.042 & 0.011 & 0.001 & 0.000 & 0.002 & 0.011 \\ 
  Shift=1 & 0.403 & 0.297 & 0.188 & 0.095 & 0.042 & 0.039 & 0.103 \\ 
  Shift=2 & 0.718 & 0.594 & 0.359 & 0.114 & 0.019 & 0.009 & 0.022 \\ 
  Shift=3 & 0.955 & 0.950 & 0.871 & 0.569 & 0.158 & 0.031 & 0.013 \\ 
  Shift=4 & 0.593 & 0.879 & 0.957 & 0.893 & 0.541 & 0.183 & 0.080 \\ 
  Shift=5 & 0.531 & 0.875 & 0.971 & 0.929 & 0.605 & 0.239 & 0.041 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Relative RMSEs of M-SSA GDP Component Predictors vs. Mean Benchmark (Including Pandemic).
%rRMSEs: Out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}.} 
\label{tab:rRMSE_mSSA_comp_mean2}}
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.999 & 0.980 & 0.951 & 0.925 & 0.915 & 0.926 & 0.955 \\ 
  Shift=1 & 1.057 & 1.051 & 1.028 & 0.996 & 0.977 & 0.978 & 0.985 \\ 
  Shift=2 & 1.031 & 1.035 & 1.026 & 0.996 & 0.964 & 0.952 & 0.964 \\ 
  Shift=3 & 1.022 & 1.035 & 1.042 & 1.026 & 0.992 & 0.965 & 0.953 \\ 
  Shift=4 & 1.002 & 1.023 & 1.044 & 1.046 & 1.024 & 0.999 & 0.984 \\ 
  Shift=5 & 1.010 & 1.032 & 1.054 & 1.060 & 1.043 & 1.013 & 0.983 \\  
   \hline
\end{tabular}
\end{table}

\begin{table}[htpb]
\caption{Relative RMSEs of Direct Forecasts vs. Mean Benchmark (Including Pandemic).
%Relative RMSEs of M-SSA GDP Component Predictors vs. Direct Forecasts (Including Pandemic).
%Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.} 
\label{tab:rRMSE_mSSA_comp_direct3}}
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.230 & 1.207 & 1.172 & 1.139 & 1.127 & 1.140 & 1.177 \\ 
  Shift=1 & 0.984 & 0.977 & 0.956 & 0.927 & 0.909 & 0.910 & 0.916 \\ 
  Shift=2 & 1.039 & 1.044 & 1.035 & 1.004 & 0.972 & 0.960 & 0.972 \\ 
  Shift=3 & 1.046 & 1.059 & 1.067 & 1.050 & 1.015 & 0.987 & 0.975 \\ 
  Shift=4 & 0.982 & 1.002 & 1.023 & 1.024 & 1.003 & 0.978 & 0.964 \\ 
  Shift=5 & 0.979 & 1.001 & 1.022 & 1.028 & 1.012 & 0.983 & 0.953 \\   
   \hline
\end{tabular}

>>>>

Let $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ denote the M-SSA predictor of HP-GDP at time $t$ for forecast horizon $h$. A GDP forecast can be obtained from the following regression specification\footnote{By default, the M-SSA criterion \eqref{eq:gen_stat_x} generates a standardized predictor and the regression coefficients are calibrated to match the level and scale of the forecast target optimally.}
\begin{equation}
\textrm{GDP}_{t+shift}=\beta_0(shift,h)+\beta_1(shift,h) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h) + \epsilon_t^{shift,h}.
\label{eq:GDP_M-SSA}
\end{equation}
%The purposes of the regression are twofold: first, the parameters enable affine transformations of the left-hand side variable, so that the standardized target $\textrm{GDP}_{t+shift}$ can be replaced by the original log-returns of GDP; second, the scale adjustment by $\beta_1(shift,h)$ compensates for the deflation (zero-shrinkage) of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, reflecting greater uncertainty as the forecast horizon $h$ extends. This simple transformation allows $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ to serve as a predictor for $\textrm{GDP}_{t+shift}$ even when $h\neq shift$. This flexibility enables leveraging the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, illustrated in Figure \eqref{multivar_vs_univar}, when $shift<h$.\\
The corresponding out-of-sample predictor $\widehat{\textrm{GDP}}_{t+shift}(h)$ is given by: 
\begin{equation}\label{eq:oos_pred}
\widehat{\textrm{GDP}}_{t+shift}(h):=\hat{\beta}_0(shift,h,t)+\hat{\beta}_1(shift,h,t) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h),
\end{equation}
where the regression parameters $\hat{\beta}_0(shift,h,t),\hat{\beta}_1(shift,h,t)$ are estimated recursively using data up to time point $t-1$ (expanding window). Since the M-SSA predictor is based on data up to Q4-2007, $\widehat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all $shift\geq 0$ and all $t$ larger than Q4-2007.\footnote{Equations \eqref{eq:GDP_M-SSA} and \eqref{eq:oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.} 
The associated out-of-sample forecast error is defined as
\begin{equation}\label{eq:oosfe}
\hat{\epsilon}_t^{shift,h}:={\textrm{GDP}}_{t+shift}-\widehat{\textrm{GDP}}_{t+shift}(h).
\end{equation}
%We now assume that the regression parameters $\hat{\beta}_0(shift,h,t)$ and$,\hat{\beta}_1(shift,h,t)$ in Equation \eqref{oos_pred} are up-dated quarterly. Note that 
In the empirical implementation, the M-SSA forecasts are held fixed, while the regression parameters $\hat{\beta}_0(\cdot)$ and $\hat{\beta}_1(\cdot)$ are updated each quarter using an expanding window beginning in 2008Q1. This procedure yields a sequence of optimally weighted M-SSA forecasts that adapt over time to evolving GDP dynamics.

@@

% Obviously, predicting HP-GDP (see Table \eqref{p_val_hpbip_wc}) is less challenging than predicting GDP (see Table \eqref{p_val_wc}), due to smoothness (autocorrelation) of the target when compared to GDP. 
Our results can be summarized as follows. When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead (Table \ref{tab:rRMSE_mSSA_direct_mean_without_covid8} in the appendix). %\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.}.
In contrast, the GDP predictors $\widehat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5 quarters ahead, see Table \ref{tab:rRMSE_mSSA_comp_mean_wc}. Additionally, $\widehat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$ (Table \ref{tab:rRMSE_mSSA_comp_direct_without_covid7}). Statistical significance of the GDP predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead. Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors, see Figures \ref{fig:multivar_vs_univar} and \ref{fig:mssa_hpc_financial_crisis}, in relation to forecast performance. Remarkably, these findings hold even though M-SSA was restricted to a relatively short in-sample window ending in Q4-2007, i.e., prior to the onset of the financial crisis.

\begin{table}[ht]
\caption{Relative RMSE of M-SSA GDP Predictors vs. Mean Benchmark.\\ 
M-SSA is fixed, based on data up to Q4-2007. Out-of-sample evaluation starting in Q1-2008, without pandemic.\\
Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$; $h$ refers to the forecast horizon for which $\textrm{M-SSA}_t^{\textrm{GDP}}(h)$ is optimized. 
Values $< 1$ indicate superior predictions from Eq.~ \eqref{eq:oos_pred}. Significance refers to the hypothesis $\alpha_1=0$ in Eq.~\eqref{eq:eq_gdp_pred}.
} 
\label{tab:rRMSE_mSSA_comp_mean_wc}
\centering
\begin{tabular}{llllllll}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.973 & $0.937^*$ & $0.888^*$ & $0.850^{**}$ & $0.855^{**}$ & $0.900^{**}$ & $0.956^*$ \\ 
  Shift=1 & 1.073 & 1.051 & 0.997 & $0.918^*$ & $0.870^{**}$ & $0.873^{**}$ & $0.906^{**}$ \\ 
  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & $0.940^*$ & $0.907^{**}$ & $0.907^{**}$ \\ 
  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & $0.941^*$ \\ 
  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & $0.942^*$ \\ 
  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & $0.963^*$ \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Relative RMSEs of M-SSA GDP Predictors vs. Direct Forecasts.\\ 
M-SSA is fixed, based on data up to Q4-2007. Out-of-sample evaluation starting in Q1-2008, without pandemic. 
Values below 1 indicate improvement over direct forecasting.
\label{tab:rRMSE_mSSA_comp_direct_without_covid7}}
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.128 & 1.086 & 1.030 & 0.985 & 0.992 & 1.044 & 1.109 \\ 
  Shift=1 & 1.280 & 1.254 & 1.189 & 1.096 & 1.037 & 1.041 & 1.081 \\ 
  Shift=2 & 1.173 & 1.181 & 1.166 & 1.104 & 1.031 & 0.995 & 0.995 \\ 
  Shift=3 & 1.012 & 1.037 & 1.058 & 1.040 & 0.989 & $0.949^*$ & $0.934^*$ \\ 
  Shift=4 & 0.982 & 1.016 & 1.056 & 1.063 & 1.022 & 0.974 & $0.941^*$ \\ 
  Shift=5 & 0.961 & 0.997 & 1.041 & 1.058 & 1.028 & 0.979 & $0.936^{*}$ \\ 
  \hline
\end{tabular}
\end{table}

@@

\begin{table}[ht]
\caption{AUCs of predictors against GDP at various forecast horizons.} 
\label{tab:auc}
\centering
\begin{tabular}{rrrrr}
  \hline
 & Direct forecast & Direct HP forecast & M-MSE & M-SSA \\ 
  \hline
Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
   \hline
\end{tabular}

@@

For illustration, we consider a simple approach based on regressing the M-SSA predictor (of HP-GDP) on future GDP. This predictor is intuitively appealing because the future HP-GDP, i.e., the target of M-SSA, is the low-frequency component of the future GDP target. Consequently, a strong link between M-SSA and future HP-GDP also indicates a connection with GDP, even if the statistical significance is obscured by the noise inherent in GDP. Out-of-sample performance suggests that this predictor is statistically significant at horizons exceeding one year. Moreover, designs optimized for larger forecast horizons outperform the mean benchmark at all considered forecast horizons, and the predictor exceeds the performance of direct forecasts---based on unfiltered or HP filtered indicators---at horizons longer than two quarters. An analysis of hit and false alarm rates within the ROC curve confirms that filtering improves predictive performance for forecast horizons exceeding two quarters. Multivariate approaches can further capitalize on leading indicators within the cross-sectional data. Additionally, the AUC is optimized through more precise control of the sign change rate within the M-SSA framework. Remarkably, these results hold even though the M-SSA filter was optimized on a relatively short in-sample window ending prior to the onset of the financial crisis. Finally, an examination of revision errors associated with quarterly updates of the proposed M-SSA-based GDP predictor indicates that the regression parameters, VAR model, and resulting M-SSA filters stabilize after an initial burn-in period. This stabilization enhances the model’s explainability and interpretability.

@@

\begin{table}[ht]
\caption{Significance of M-SSA GDP Component Predictor (Including Pandemic).\\
p-values for $H_0: {\alpha_1} = 0$, HAC-adjusted Wald test, Equation \eqref{eq:eq_gdp_pred}.\\ including the pandemic.} 
\label{tab:p_val1}
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.127 & 0.064 & 0.019 & 0.003 & 0.001 & 0.002 & 0.012 \\ 
  Shift=1 & 0.439 & 0.314 & 0.194 & 0.093 & 0.033 & 0.017 & 0.036 \\ 
  Shift=2 & 0.718 & 0.565 & 0.326 & 0.098 & 0.020 & 0.008 & 0.010 \\ 
  Shift=3 & 0.946 & 0.975 & 0.925 & 0.657 & 0.222 & 0.059 & 0.028 \\ 
  Shift=4 & 0.327 & 0.699 & 0.964 & 0.942 & 0.619 & 0.237 & 0.102 \\ 
  Shift=5 & 0.383 & 0.755 & 0.946 & 0.910 & 0.581 & 0.241 & 0.055 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Relative RMSEs of M-SSA GDP Component Predictors vs. Mean Benchmark (Including Pandemic).
%rRMSEs: Out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}.} 
\label{tab:rRMSE_mSSA_comp_mean2}}
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.005 & 0.986 & 0.960 & 0.935 & 0.927 & 0.939 & 0.964 \\ 
  Shift=1 & 1.060 & 1.054 & 1.032 & 0.998 & 0.973 & 0.965 & 0.970 \\ 
  Shift=2 & 1.028 & 1.029 & 1.020 & 0.993 & 0.963 & 0.950 & 0.952 \\ 
  Shift=3 & 1.018 & 1.027 & 1.034 & 1.022 & 0.994 & 0.972 & 0.961 \\ 
  Shift=4 & 0.990 & 1.006 & 1.025 & 1.032 & 1.017 & 0.998 & 0.985 \\ 
  Shift=5 & 1.000 & 1.016 & 1.037 & 1.046 & 1.032 & 1.008 & 0.985 \\  
   \hline
\end{tabular}
\end{table}

\begin{table}[htpb]
\caption{Relative RMSEs of of M-SSA GDP Component Predictors  vs. Direct Forecasts (Including Pandemic).
%Relative RMSEs of M-SSA GDP Component Predictors vs. Direct Forecasts (Including Pandemic).
%Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.} 
\label{tab:rRMSE_mSSA_comp_direct3}}
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.237 & 1.215 & 1.183 & 1.152 & 1.141 & 1.156 & 1.187 \\ 
  Shift=1 & 0.986 & 0.980 & 0.960 & 0.929 & 0.905 & 0.898 & 0.903 \\ 
  Shift=2 & 1.036 & 1.037 & 1.029 & 1.001 & 0.971 & 0.957 & 0.960 \\ 
  Shift=3 & 1.042 & 1.051 & 1.058 & 1.045 & 1.018 & 0.995 & 0.983 \\ 
  Shift=4 & 0.970 & 0.985 & 1.004 & 1.010 & 0.996 & 0.978 & 0.965 \\ 
  Shift=5 & 0.970 & 0.986 & 1.007 & 1.015 & 1.002 & 0.978 & 0.956 \\   
   \hline
\end{tabular}

### 2025-08-16 Sweave/BIP_predictor.Rnw
up

select_vec_multi<-c("BIP","ip","ifo_c","ESI","spr_10y_3m")
x_mat<-data[,select_vec_multi]
tail(x_mat)
rownames(x_mat)<-rownames(data)
n<-dim(x_mat)[2]
# Number of observations
len<-dim(x_mat)[1]

@@

read_obj<-read_data_func_monthly(path.data,select_vec,h0,trim_threshold,na_rw_lin,data_file_name,diff_lag)
data_table<-read_obj$data_select_unshifted_xts
colnames(data_table)[1]<-"GDP"
tail(data_table,20)

>>>>

select_vec_multi<-c("BIP","ip","ifo_c","ESI","spr_10y_3m")
x_mat<-data[,select_vec_multi]
head(x_mat)
tail(x_mat)
rownames(x_mat)<-rownames(data)
n<-dim(x_mat)[2]
# Number of observations
len<-dim(x_mat)[1]

@@

read_obj<-read_data_func_monthly(path.data,select_vec,h0,trim_threshold,na_rw_lin,data_file_name,diff_lag)
data_table<-read_obj$data_select_unshifted_xts
colnames(data_table)[1]<-"GDP"
head(data_table,20)
tail(data_table,20)
